---
title: "Final Project_R"
author: "Korey Bernhardt"
date: "12/11/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries

```{r}
library(dplyr)
library(ggformula)
library(readr)
```

**Read data**

Pull in the data created from the Python program, and check the first few columns to make sure it's what's expected.

```{r}
tweets = read.csv("unique_tweets.csv",header=TRUE, sep=",",)
head(tweets)
```

**Adjust data and add fields**

In order to make the data a little more  user friendly, I changed the Sentiment column to hold Fan or Foe if the sentiment is positive or negative, respectively.  I also added a column to hold if the location has an NFL team as a Yes or No value based on the number of teams the location has.  I want both ways available to analyze the data.

```{r}
tweets = 
  tweets %>%
  mutate(HasNFLTeam = ifelse(team_count > 0,'Yes','No' ))%>%
  mutate(FanSentiment = case_when(
    sentiment == "positive" ~ "Fan",
    sentiment == "negative" ~ "Foe",
    TRUE ~ "Neutral"))
```

**Charting tweet volumen by location**

I limited this to the top 10 locations by tweet volume to make it easier to read. 

```{r}
tweets_count_location =
  tweets%>%
  group_by(FanSentiment,location)%>%
  summarise(location_count = n())%>%
  top_n(10, location_count)%>%
  mutate(location = reorder(location,-location_count))%>%
  gf_col(location_count ~ location, fill=~FanSentiment,"dark green")%>%
  gf_refine(scale_fill_manual(values = c("dark green","gold")))%>%
  gf_labs(title = "Tweet Volume by State, Top 10", y="# Tweets", x = "State")

tweets_count_location

```
**Charting Categories**

I charted two ways of looking at things.  One based on state and whether the state had a NFL team.  The other based on state and how many NFL teams the state has. 

```{r}
tweets_by_location_hasteam =
  tweets%>%
  count(location, FanSentiment, HasNFLTeam, sort = TRUE)%>%
  gf_col(n ~ HasNFLTeam,fill=~FanSentiment, position = 'dodge')%>%
  gf_refine(scale_fill_manual(values = c("dark green","gold")))%>%
  gf_labs(title = "Tweet Volume and Sentiment by NFL Team Presence", 
          y="# Tweets", x = "NFL Team Presence?")
  

tweets_by_location_teamcount =
  tweets%>%
  count(location, FanSentiment, team_count, sort = TRUE)%>%
  gf_col(n ~ team_count,fill=~FanSentiment,position = 'dodge')%>%
  gf_refine(scale_fill_manual(values = c("dark green","gold")))%>%
  gf_labs(title = "Tweet Volume and Sentiment by NFL Team Count", 
          y="# Tweets", x = "Number of NFL Teams")

tweets_by_location_hasteam   
tweets_by_location_teamcount

```

**Hypothesis Test**

The question I planned to answer was where are packer fans outside of Wisconsin located? I predicted they would be more heavily prevalent in locations that don’t have NFL teams as there is less of a “default” affiliation with a specific team.

Null Hypotheses: There is no relationship between tweet sentiment for the packers and states having an NFL team.

Alternative Hypothesis: There is a relationship between tweet sentiment for the packers and states having an NFL team.

Using the chi-square test of independence, based on the data we have the null hypothesis cannot be rejected, and therefore there does not appear to be a relationship between tweet sentiment and states having an NFL team.

```{r}
chisq.test (tweets$FanSentiment, tweets$HasNFLTeam)
chisq.test (tweets$FanSentiment, tweets$team_count)

```
```{r}
write.csv(tweets,file="unique_tweets_R_updated.csv",row.names=FALSE,na="")
```

